# Reproducible Research: Peer Assessment 1

# Activity Monitoring Data Analysis
========================================================

## Loading and preprocessing the data

### Load activity.csv file into dataframe filedata, and create stesdate dataframe with date and total number of steps on each day.

```{r}
filedata<-read.csv("activity.csv")
stepsdate<-tapply(filedata$steps,filedata$date,sum)
```

## What is mean total number of steps taken per day?

### Histogram showing the total number of steps taken each day versus date:

```{r}
hist(stepsdate,xlab="total steps per day",ylab="number of days",main="histogram of daily steps")
```

### Find mean step per day by applying mean function to steps on each date:
```{r}
meansteps<-aggregate(filedata$steps,by=list(filedata$date),FUN=mean)
names(meansteps)<-c("date","meanstep")
kable(meansteps)
```

### Find median step per day by applying median function to stpes on each date, with not valid number ignored:
```{r}
filedatawona<-subset(filedata,!is.na(filedata$steps))
mediansteps<-aggregate(filedatawona$steps,by=list(filedatawona$date),FUN=median)
names(mediansteps)<-c("date","median")
kable(mediansteps)
```

## What is the average daily activity pattern?

### Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis):
```{r}
stepsdata<-subset(filedata,!is.na(filedata$steps))
stepts<-aggregate(stepsdata$steps, by=list(stepsdata$interval), FUN=mean)
plot(stepts,type="l",xlab="Interval", ylab="Average Steps")
```


```{r}
names(stepts)<-c("interval","avgsteps")
maxstepinterval<-subset(stepts,stepts$avgsteps==max(stepts$avgsteps))$interval
```

### Interval `r maxstepinterval` contains the maximum number of steps.

## Imputing missing values

```{r}
numofnas<-length(subset(filedata,is.na(filedata$steps))[,1])
```

### There are `r numofnas` NAs on the file.

### Using the mean steps of corresponding 5-minute interval to fill in missing data:

```{r}
fulldata<-filedata
for(n in 1:length(fulldata[,1]))
{
    record<-fulldata[n,]
    if(is.na(record$steps))
    {
      interval<-record$interval
      meanstep<- stepts[stepts$interval==interval,]$avgsteps
      fulldata[n,]$steps = meanstep
    }
}

fullstepsdate<-tapply(fulldata$steps,fulldata$date,sum)
```

### Histogram of the total number of steps taken each day with missing value filled in:
```{r}
hist(fullstepsdate,xlab="total steps per day",ylab="number of days",main="histogram of daily steps")
```

### Find mean step per day by applying mean function to steps on each date with missing value filled in:
```{r}
fullmeansteps<-aggregate(fulldata$steps,by=list(fulldata$date),FUN=mean)
names(fullmeansteps)<-c("date","meanstep")
kable(fullmeansteps)
```

### Find median step per day by applying median function to stpes on each date with missing value filled in:
```{r}
mediansteps<-aggregate(fulldata$steps,by=list(fulldata$date),FUN=median)
names(mediansteps)<-c("date","median")
kable(mediansteps)
```

## Are there differences in activity patterns between weekdays and weekends?

### plot of the daily patterns for weekdays and weekends:
```{r}
days<-character()
for(n in 1:length(fulldata[,1]))
{
   dt<-fulldata[n,]$date
   dt<-as.Date(dt,"%Y-%m-%d")
   if(weekdays(dt) %in% c("Sunday","Saturday"))
     days<-c(days,"weekend")
   else
     days<-c(days,"weekday")
}
days<-factor(days)
datawday<-cbind(fulldata,days) 

dataweekday<-subset(datawday,datawday$days=="weekday")
dataweekend<-subset(datawday,datawday$days=="weekend")

stepwdts<-aggregate(dataweekday$steps, by=list(dataweekday$interval), FUN=mean)
names(stepwdts)<-c("interval","avg. steps")

stepwendts<-aggregate(dataweekend$steps, by=list(dataweekend$interval), FUN=mean)
names(stepwendts)<-c("interval","avg. steps")

par(mfrow=c(2,1))
plot(stepwdts,type="l",xlab="Interval", ylab="Average Steps",main="weekdays")
plot(stepwendts,type="l",xlab="Interval", ylab="Average Steps",main="weekends")
```

### Above panel plot shows differences between weekday activities and weekend activities. People are more active in the morning on weekdays, and on weekends their activities are likely in the afternoon.